### 2018-04-06 ident/docs/draft_2.tex
change draft file name from 1 to 2

\begin{document}
	\begin{center}
		\begin{Large}
			Practical Identification and Experimental Design for Parameter Estimation in Kinetic Models of Metabolism
		\end{Large}\\
		Shyam Srinivasan\textsuperscript{a}, William R. Cluett\textsuperscript{a} and Radhakrishnan Mahadevan\textsuperscript{*,a,b}\\
	\end{center}
	a - Department of Chemical Engineering and Applied Chemistry, University of Toronto, Toronto, ON, Canada.\\
	b - Institute for Biomaterials and Biomedical Engineering, University of Toronto, Toronto, ON, Canada.\\
	{*} Corresponding author
	\section*{Abstract}
	\section{Introduction:}
	The use of metabolic engineering spans a wide variety of applications. Some notable examples include the design of microorganisms for the biosynthesis of commodity and specialty chemicals \parencite{Andreozzi2016}, engineering mammalian cells as therapeutic targets for cures to some ailments affecting humans \parencite{DiFilippo2016,Apaolaza2017}, and changing the constituents of the human gut microbial community to cure related diseases \parencite{Zerfab2018}. These applications require us to understand the numerous complex interaction, their roles in cell function, and sometimes even the mechanisms behind these interactions \parencite{Pushalkar2018}. Computational models offer a systematic way to integrate available experimental data, and to study and understand these interactions through mathematical representations of the biological systems in which these interactions occur \parencite{Bordbar2014a,Saa2017}. They are also used to predict changes in cell function based on changes in the type and nature of the modeled interactions \parencite{Andreozzi2016}, or aid in the identification of therapeutic targets for drug discovery and development \parencite{Bordbar2015,Chandrasekaran2017}

	Constraint-based models (CBMs) of metabolism are one such category of models that are used to improve our understanding of metabolism by representing it as a stoichiometric network of reactions that operate under a pseudo-steady state assumption \parencite{Bordbar2014a}. The ability of CBMs to shine light on the nonintuitive interactions that govern cellular metabolism is leveraged to engineer and asses the impact of designs that alter the ability of a cell to grow, or produce a desired metabolite \parencite{Maia2016}. However, in CBMs, since metabolism is assumed to operate under a pseudo steady state, the metabolite concentrations within the metabolic network are assumed to be constant, and changes in metabolite concentrations are not modeled. Furthermore, since CBMs represent metabolism using only the stoichiometry of its constituent reactions, they do not account for the various non-catalytic regulatory interactions that are also responsible of metabolic function. These shortcomings prevent CBMs from being used to fully understand the steady state as well as the dynamic characteristics of metabolic networks.

	Recently, the focus has shifted towards understanding the implications of metabolite concentrations and associated regulatory interactions on the steady state and the dynamic characteristics of metabolism \parencite{Saa2017}. Kinetic models accounting for the thermodynamic and regulatory interactions that underly metabolic networks in addition to its stoichiometry are used for this purpose \parencite{Link2014}. These models can not only help us learn about previously unknown characteristics of metabolism \parencite{Kotte2014}, but can also improve predictions about the impact of engineering design perturbations on metabolism, and propose alternative designs to achieve various goals \parencite{Andreozzi2016,Khodayari2016}.

	Kinetic models differ from CBMs in their use of heavily parameterized mechanistic enzyme kinetic rate laws to model enzyme catalyzed fluxes within a metabolic network. These parameters represent various aspects of the enzyme kinetic rate laws \parencite{Srinivasan2015,Saa2017}. Hence, the use of kinetic models requires information on all the enzyme kinetic rate laws that will be used to model all the fluxes within a metabolic network, as well as numerical values for all the parameters used in these rate laws. Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where kinetic rate laws and parameter values might play a crucial role \parencite{Vital-Lopez2006,Srinivasan2017}.

>>>>

\begin{document}
	\begin{center}
		\begin{Large}
			Practical Identification and Experimental Design for Parameter Estimation in Kinetic Models of Metabolism
		\end{Large}\\
		Shyam Srinivasan\textsuperscript{a}, William R. Cluett\textsuperscript{a} and Radhakrishnan Mahadevan\textsuperscript{*,a,b}\\
	\end{center}
	a - Department of Chemical Engineering and Applied Chemistry, University of Toronto, Toronto, ON, Canada.\\
	b - Institute for Biomaterials and Biomedical Engineering, University of Toronto, Toronto, ON, Canada.\\
	{*} Corresponding author
	\section*{Abstract}
	\section{Introduction:}
	The use of metabolic engineering spans a wide variety of applications. Some notable examples include the design of microorganisms for the biosynthesis of commodity and specialty chemicals \parencite{Andreozzi2016}, engineering mammalian cells as therapeutic targets for cures to some ailments affecting humans \parencite{DiFilippo2016,Apaolaza2017}, and changing the constituents of the human gut microbial community to cure related diseases \parencite{Zerfab2018}. These applications require us to understand the numerous complex interaction, their roles in cell function, and sometimes even the mechanisms behind these interactions. Computational models offer a systematic way to integrate available experimental data, and to study and understand these interactions through mathematical representations of the biological systems in which these interactions occur \parencite{Bordbar2014a,Saa2017}. They are also used to predict changes in cell function based on changes in the type and nature of the modeled interactions \parencite{Andreozzi2016}, or aid in the identification of therapeutic targets for drug discovery and development \parencite{Bordbar2015,Chandrasekaran2017}

	Constraint-based models (CBMs) of metabolism are used to improve our understanding of metabolism by representing it as a stoichiometric network of reactions that operate under a pseudo-steady state assumption \parencite{Bordbar2014a}. The ability of CBMs to shine light on the nonintuitive interactions that govern cellular metabolism is leveraged to engineer and asses the impact of designs that alter the ability of a cell to grow, or produce a desired metabolite \parencite{Maia2016}. However, in CBMs, metabolism is assumed to operate under a pseudo steady state. Consequently, the metabolite concentrations within the metabolic network are assumed to be constant, and changes in metabolite concentrations are not modeled. Furthermore, since CBMs represent metabolism using only the stoichiometry of its constituent reactions, they do not account for the various non-catalytic regulatory interactions that are also responsible of metabolic function. These shortcomings prevent CBMs from being used to fully understand the steady state as well as the dynamic characteristics of metabolic networks.

	In contrast, the implications of regulatory interactions and changes in metabolite concentrations on different characteristics of metabolism can be studied using kinetic models of metabolism \parencite{Saa2017}. These models account for the thermodynamic and regulatory interactions that underly metabolic networks in addition to its stoichiometry \parencite{Link2014}. Kinetic models can not only help us better understand lesser known and understood characteristics of metabolism like bistability \parencite{Kotte2014}, and their role in human health, but can also improve predictions about the impact of engineering design perturbations on metabolism, and propose alternative designs to achieve metabolite production goals \parencite{Khodayari2016}.

	Kinetic models differ from CBMs in their use of heavily parameterized mechanistic enzyme kinetic rate laws to model enzyme catalyzed fluxes within a metabolic network. These parameters represent various aspects of the enzyme kinetic rate laws \parencite{Srinivasan2015,Saa2017}. Hence, the use of kinetic models requires information on all the enzyme kinetic rate laws that will be used to model all the fluxes within a metabolic network, as well as numerical values for all the parameters used in these rate laws. Analyzing the ability of a metabolic network to exhibit dynamic characteristics like multiple steady states and oscillations, irrespective of the structure of the network, is one example where kinetic rate laws and parameter values might play a crucial role \parencite{Srinivasan2017}.
